Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-06-07 doc/invt.tex
small changes

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014}.
%
We define a bin-average-deducted bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  ,
\label{eq:x_def}
\end{equation}
%
where the bin average, $\bav{v}(t)$, is defined as
%
\begin{equation}
  \bav{v}(t)
  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vav_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$
(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
%
The total error is given by
%
\begin{equation}
  \Err(t)
  =
  \sum_{i = 1}^n \rho_i \, \left\langle x_i^2(t) \right\rangle
  .
\label{eq:error_def}
\end{equation}

%Although the above error weighting distribution, $\pmb\rho$,
%is usually flat,
%i.e. $\rho_i \equiv 1/n$,
%in practice,
%a non-flat distribution
%is sometimes more convenient in analyzing
%some multiple-bin updating schemes later on.
%
Note that the distribution, $\pmb\rho$, is arbitrary
and independent of
the target distribution,
$\mathbf p = (p_1, \dots, p_n)$,
although it is usually flat.\footnote{In
well-tempered metadynamics,
$\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
  barducci2008, dama2014}:
$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
whereas $\pmb\rho$ is flat.}
%
From Eqs. \eqref{eq:x_def} and \eqref{eq:vav_def},
we have
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, x_i(t) = 0
  .
\label{eq:rhox_sum}
\end{equation}
%

@@

To proceed, we will
approximate the finite difference Eq. \eqref{eq:wl_update}
by a differential equation\footnote{This
approximation requires us to shift the origin of time by $1$ in some cases.}
%
\begin{equation}
  \dot u_i(t)
  \equiv
  \frac{ d u_i(t) } { dt }
  =
  \alpha(t) \, \frac{ h_i(t) } { p_i }
  \equiv
  \alpha(t) \, \eta_i(t)
  ,
  \label{eq:vt_diffeq}
\end{equation}
%
with
%
$
h_i(t) = \delta_{i, i(t)}
$
%\begin{equation}
%  h_i(t) = \delta_{i, i(t)}
%  ,
%  \label{eq:h_def}
%\end{equation}
%
being the instantaneous histogram,
which is equal to $1.0$
for the bin $i(t)$ at time $t$
or zero otherwise,
and we have defined a relative histogram
$\eta_i(t) \equiv h_i(t) /p_i$
in the last step.
%
The shifted bias potential, $v_i(t)$,
follows the same differential equation
as the shift is a constant.

We now try to split the histogram into a deterministic averaged part
and a random fluctuation part, $\zeta_i(t)$,
%
\begin{equation}
  \eta_i(t) = \langle \eta_i(t) \rangle + \zeta_i(t).
  \label{eq:h_split}
\end{equation}
%
The deterministic part can be related
to an ``ensemble average'',
where the ensemble consists of many similar simulation copies
sharing the same schedule $\alpha(t)$
and the same bias potential at time $t$.
%
The initial states and the stochastic forces
during the process may differ, however.

For sufficiently small $\alpha(t)$,
the bias potential remains roughly the same for a short period.
%
Then,
the sampling process is approximately Markovian, % of a finite order,
and we may assume Eq. \eqref{eq:pi_p_V}
for the deterministic part
%
\begin{align}
  \langle \eta_i(t) \rangle
  %&=
  %\frac{ \langle h_i(t) } { p_i } \rangle
  &\approx
  \frac{ \pi_i(t) } { p_i }
  =
  \frac{                       e^{-x_i(t)} }
       { \sum_{j = 1}^n p_j \, e^{-x_j(t)} }
  \notag
  \\
  &\approx
  1 - x_i(t) + \sum_{j = 1}^n p_j \, x_j(t)
  ,
  \notag
  %\label{eq:h_ave}
\end{align}
%
where we have assumed the smallness
of $x_i(t)$ in the linear expansion
and used $\sum_{j=1}^n p_j = 1$
on the second line.
%
\note{
The second step follows from
$$
\frac{ \langle h_i(t) \rangle }
     { p_i }
\approx
\frac{                       1 - x_i  }
     { \sum_{ r = 1 }^n p_j (1 - x_j) }
=
\frac{                       1 - x_i  }
     { 1 - \sum_{ j = 1 }^n p_j \, x_j }
.
$$
}%
%
Deducting the bin average [similar to Eq. \eqref{eq:vav_def}],
we have
%
\begin{align}
  \langle \eta_i(t)  - \eta_\mathrm{av}(t) \rangle
  \equiv
  \langle \eta_i(t) \rangle -
  \sum_{j=1}^n \rho_j \langle \eta_j(t) \rangle
  &\approx
  - x_i(t)
  .
  \label{eq:sh_ave}
\end{align}

%
For the histogram fluctuation part,
the average is zero by definition,
$\langle \zeta_i(t) \rangle = 0$,
%
and we will approximate it as white noise.
%i.e.
%%The noise is not necessarily white,
%%and the autocorrelation function
%%depend only on the time difference
%%
%\begin{equation}
%  \left\langle
%    \zeta_i(t) \, \zeta_j(t')
%  \right\rangle
%  =
%  %\sigma_{ij}(t - t')
%  G^{(0)}_{ij} \, \delta(t - t')
%  ,
%  \notag
%  %\label{eq:zeta_zeta_correlation}
%\end{equation}
%%
%for some value of $G^{(0)}_{ij}$,
%where $\sigma_{ij}(t)$ is an even function
%that vanishes at large $t$.
%
This approximation is valid if the simulation period, $T$,
is much longer than the autocorrelation time of noise.
%
The bin-average-deducted fluctuation
%
\begin{equation}
  \xi_i(t)
  \equiv
  \zeta_i(t) - \zeta_\mathrm{av}(t)
  \equiv
  \zeta_i(t)
  -
  \sum_{ j = 1 }^n
    \rho_j \, \zeta_j(t)
  ,
\label{eq:xi_def}
\end{equation}
%
would also be governed by white noise as is $\zeta_i(t)$:
%i.e. we may write
\begin{equation}
  \langle \xi_i(t) \, \xi_i(t') \rangle
  = G_i \, \delta(t - t')
  ,
  \label{eq:G_def}
\end{equation}
for some $G_i$,
and $\delta(t)$ is the Dirac's delta function.
%
For non-white noise,
the equivalent value of $G_i$ can be approximately obtained as
$\sum_{t = -\infty}^\infty \langle \xi_i(t) \, \xi_i(0) \rangle$.
%
\note{More explicitly,
if the transition matrix of a period, $\tau$,
is $T^\tau_{ij}$,
then
$
  p_i p_j \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  \langle h_i(t) \, h_j(t') \rangle
  -
  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
  =
  T^{t - t'}_{ij} \, \pi_j - \pi_i \, \pi_j.
$
}%

From Eqs.
\eqref{eq:x_def}, \eqref{eq:vt_diffeq}-\eqref{eq:xi_def},
%\eqref{eq:v_def},
%\eqref{eq:x_def},
%\eqref{eq:vav_def},
%\eqref{eq:rhox_sum},
%\eqref{eq:vt_diffeq},
%\eqref{eq:h_split}
%\eqref{eq:h_ave},
%\eqref{eq:dvbardt},
we get % a set of decoupled equations
%
\begin{equation}
  \dot x_i(t)
  =
  \alpha(t) \, \left[ \eta_i(t) - \eta_\mathrm{av} \right]
  \approx
  -\alpha(t) \, \left[ x_i(t) - \xi_i(t) \right]
  ,
  \notag
  %\label{eq:dxdt_singlebin}
\end{equation}
%
and the formal solution is % of Eq. \eqref{eq:dxdt_singlebin} is
%
\begin{equation}
x_i(T)
=
x_i(0) \, e^{-q(T)}
+
\int_0^T
  \dot y\bigl( q(t) \bigr) \, \xi_i(t) \, dt,
\notag
%\label{eq:xt_solution}
\end{equation}
%
where
%
$
q(t) \equiv \int_0^t \alpha(t') \, dt',
$
%
and
%
\begin{align}
y(q')
&\equiv
e^{q' - q(T)}.
\label{eq:y_def}
\end{align}

@@

The error depends implicitly on the curve, $\alpha(t)$,
or, equivalently, the curve, $q(t)$,
via $x_i(T)$, and
our aim is to find the $q(t)$
that minimizes this error.
%
We now fix the endpoint value, $q(T)$,
and vary the curve, $q(t)$ ($0 < t < T$).
%
This is equivalent to varying the curve $y(q(t))$
with fixed endpoints,
$y\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$y\bigl( q(T) \bigr) = 1$.
%
Note also that the first term on the right-hand side
of Eq. \eqref{eq:ET_average} is fixed during
the process,
while the second term is the action of a free particle.
Thus, the velocity of $y(q(t))$ should be a constant:
%
%We will further demand that
%%
%\begin{equation}
%  \lim_{t \to \infty} q(t) \to \infty.
%  \label{eq:qt_limit}
%\end{equation}
%%
%Then, for a large $t$,
%the second term on the right-hand side
%of Eq. \eqref{eq:xt_solution} can be neglected, and
%%
%\begin{align}
%\left\langle x^2(t) \right\rangle
%%&=
%%\int_0^t \int_0^t \dot y(t') \, \dot y(t'')
%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%%\notag
%%\\
%&=
%\int_0^t \int_0^t
%  \dot y\bigl( q(t') \bigr) \,
%  \dot y\bigl( q(t'') \bigr) \,
%  \kappa(t' - t'') \, dt'' \, dt'.
%\label{eq:x2t_average_asym}
%\end{align}
%
%By the Euler-Lagrange equation, we get
%$$
%\begin{aligned}
%0
%&=
%\frac{d}{dt} \int_0^T
%  \dot y\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
%\\
%&= \int_0^T
%  \ddot y\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
%\end{aligned}
%$$
%where we have dropped the boundary terms
%by assuming Eq. \eqref{eq:kappa_limit}
%and a large $T$.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t y^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
%In order for this to hold for any $t$,
%we must have
%%
%$
%\ddot y\bigl( q(t') \bigr) = 0
%,
%$
%or
\begin{equation}
  \dot y\bigl( q(t) \bigr) = c
  .
\label{eq:dydt_const}
\end{equation}
%
Then, by using Eq. \eqref{eq:y_def}, we have
%
\begin{equation}
  e^{ q(t) - q(T) }
  =
  c \, (t + t_0)
  =
  \frac{ t + t_0 }
       { T + t_0 }
  ,
\label{eq:expqt}
\end{equation}
where $t_0$ is some constant,
and the other constant, $c$, has been determined
from the $t = T$ case
in the second step.
%
Taking the logarithm,
and differentiating both sides
with respect to $t$,
we get the inverse-time schedule
%
\begin{equation}
  \alpha(t) = \frac{ 1 }{ t + t_0 }
  ,
\label{eq:alpha_invt1}
\end{equation}
%
and Eq. \eqref{eq:alpha_invt}
is the special case of $t_0 = 0$.
%

@@

The matrix, $\mathbf w$, is subject to three conditions.
%
First, we have a fixed-point condition\cite{bussi2006, dama2014}.
%
To sample the desired distribution,
$\mathbf p = (p_1, \dots, p_n)$,
${\dot u}_i(t)$
in Eq. \eqref{eq:ut_diffeq_mbin}
should be a constant of $i$
if $h_j(t)$ were the same as $p_j$,
i.e.
$\sum_{j=1}^n w_{ij}$ should be a constant
to allow the bias potential to grow uniformly
in the asymptotic regime.
%
By a proper overall scaling of $\mathbf w$,
we may write this condition as
%
\begin{equation}
  \sum_{j = 1}^n w_{ij} = 1
  .
\label{eq:w_sumj}
\end{equation}
%
In other words, $(1, \dots, 1)^T$
is a right eigenvector of $\mathbf w$
with eigenvalue $1$.
%
Thus, the transpose $\mathbf w^T$
resembles a transition matrix,
except that some elements can be negative
in our cases.
%
Below we will take advantage of the analogy
and borrow techniques used
in studying transition matrices\cite{vankampen}
(we will only use results that are unaffected by the allowance
of negative elements into $\mathbf w$).

@@

We can further define
a generalized Fourier transform, $\mathcal{F}$,
for variable, $A_i$,
%
\begin{equation}
  {\tilde A}_k
  \equiv \mathcal{F}[A_i]_k
  \equiv \sum_{i = 1}^n \phi_{ki} \, A_i
  ,
  \label{eq:gft_def}
\end{equation}
%
and the inverse transform is
%
\begin{equation}
  A_i =
  \frac{1}{\rho_i} \sum_{k = 1}^n \phi_{ki} \, \tilde{A}_k
  .
  \label{eq:gft_inv}
\end{equation}
%
Using this definition,
we can rewrite Eq. \eqref{eq:ortho1} as
%
\begin{equation}
  \mathcal F[1]_k = \delta_{k0}
  ,
  \label{eq:ortho0}
\end{equation}
%
and the bin average, Eq. \eqref{eq:vav_def} can be written as
\begin{equation}
  v_\mathrm{av} = \mathcal F[v_i]_0 = \tilde v_0
  .
  \label{eq:vav_vt0}
\end{equation}

We can now diagonalize
the multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
as
%
$$
{\tilde v}_k(t + 1) =
{\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
{\tilde \eta}_k(t)
,
$$
or in the continuous-time form,
%
\begin{equation}
  \dot{\tilde v}_k(t)
  =
  \alpha(t) \, \lambda_k \, {\tilde \eta}_k(t)
  .
  \label{eq:ut_diffeq_mbin}
\end{equation}
%
Deducting the bin average, we get
%
$$
\begin{aligned}
  \dot {\tilde x}_k(t)
  &= \alpha(t) \, \lambda_k
  \left[
    {\tilde \eta}_k(t)
    -
    {\tilde \eta}_\mathrm{av}(t) \, \delta_{k0}
  \right]
  \\
  &\approx
  -\alpha(t) \, \lambda_k \,
  \bigl[ {\tilde x}_k(t) - {\tilde \xi}_k (t) \bigr]
  ,
\end{aligned}
$$
where we have used the linear approximation
introduced in Sec. \ref{sec:sbin_diffeq}.

@@

The error defined in Eq. \eqref{eq:error_def}
can be rewritten in terms of the eigenmodes of $\mathbf w$,
%
\begin{align}
  \Err
  \equiv
  \sum_{i = 1}^n \rho_i \,
    \left\langle x_i^2 \right\rangle
  =
  \sum_{i, j = 1}^n \sum_{k=0}^{n-1}
   \phi_{ki} \, \phi_{kj} \,
    \left\langle x_i \, x_j \right\rangle
  =
  \sum_{k = 0}^{n - 1}
    \left\langle
      {\tilde x}_k^2
    \right\rangle
  ,
\notag
\end{align}
%
where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
%
From Eq. \eqref{eq:eigenmode0},
the first eigenmode,
${\tilde x}_0 = \sum_{j=1}^n \rho_j \, x_j$,
vanishes identically by Eq. \eqref{eq:rhox_sum}.
%
Thus, we can start the last sum %in Eq. \eqref{eq:xt2_sum}
from $k = 1$,
%
\begin{equation}
  \Err(T)
  =
  \sum_{k = 1}^{n - 1}
    \left\langle
      {\tilde x}_k^2(T)
    \right\rangle
  .
\label{eq:xt2_sum}
\end{equation}

>>>>

Second, the bias potential following Eq. \eqref{eq:wl_update}
generally grows over time.
%
Since a uniform increase of $u_i(t)$, or, equivalently, that of $v_i(t)$,
does not affect the resulting distribution, $\pi_i(t)$,
the real deviation of the bias potential
depends only on the difference between $v_i(t)$
and a baseline value, $\bav{v}(t)$,
for average growth\cite{
dama2014}.
%
We define a bin-average-deducted bias potential as
%
\begin{equation}
  v_{*i}(t) \equiv v_i(t) - \bav{v}(t)
  ,
\label{eq:x_def}
\end{equation}
%
where the bin average, $\bav{v}(t)$, is defined as
%
\begin{equation}
  \bav{v}(t)
  \equiv \sum_{i = 1}^n \rho_i \, v_i(t)
  ,
\label{eq:vav_def}
\end{equation}
for some distribution,
$\pmb \rho = (\rho_1, \dots, \rho_n)$
(with $\rho_i \ge 0$ and $\sum_{i = 1}^n \rho_i = 1$).
%
The total error is given by
%
\begin{equation}
  \Err(t)
  =
  \sum_{i = 1}^n \rho_i \,
  \left\langle v_{*i}^2(t) \right\rangle
  .
\label{eq:error_def}
\end{equation}

%Although the above error weighting distribution, $\pmb\rho$,
%is usually flat,
%i.e. $\rho_i \equiv 1/n$,
%in practice,
%a non-flat distribution
%is sometimes more convenient in analyzing
%some multiple-bin updating schemes later on.
%
Note that the distribution, $\pmb\rho$, is arbitrary
and independent of
the target distribution,
$\mathbf p = (p_1, \dots, p_n)$,
although it is usually flat.\footnote{In
well-tempered metadynamics,
$\mathbf p$ is an attenuated version of
$\mathbf p^*$\cite{
  barducci2008, dama2014}:
$\ln p_i = \ln {p^*_i}/(1+\Delta) + \mathrm{const.}$ for some positive $\Delta$,
whereas $\pmb\rho$ is flat.}
%
%From Eqs. \eqref{eq:x_def} and \eqref{eq:vav_def},
%we have
%
%\begin{equation}
%  \sum_{i = 1}^n \rho_i \, v_{*i}(t) = 0
%  .
%\label{eq:rhox_sum}
%\end{equation}
%

@@

To proceed, we will
approximate the finite difference Eq. \eqref{eq:wl_update}
by a differential equation\footnote{This
approximation requires us to shift the origin of time by $1$ in some cases.}
%
\begin{equation}
  \dot u_i(t)
  \equiv
  \frac{ d u_i(t) } { dt }
  \approx
  \alpha(t) \, \frac{ h_i(t) } { p_i }
  \equiv
  \alpha(t) \, \eta_i(t)
  ,
  \label{eq:ut_diffeq}
\end{equation}
%
with
%
$h_i(t) = \delta_{i, i(t)}$
%\begin{equation}
%  h_i(t) = \delta_{i, i(t)}
%  ,
%  \label{eq:h_def}
%\end{equation}
%
being the instantaneous histogram,
which is equal to $1.0$
for the bin $i(t)$ at time $t$
or zero otherwise,
and we have defined a relative histogram
$\eta_i(t) \equiv h_i(t) /p_i$
in the last step.
%
The shifted bias potential, $v_i(t)$,
follows the same differential equation
as the shift is a constant.
%
Thus, we have
%
\begin{equation}
  \dot v_{*i}(t)
  \approx
  \alpha(t) \, \eta_{*i}(t)
  ,
  \label{eq:vt_diffeq}
\end{equation}
%
where we have defined the bin-average-deducted
histogram, $\eta_{*i}(t)$, in a similar way to
$v_{*i}(t)$ [cf. Eq. \eqref{eq:x_def}].

We now split the histogram into
a deterministic averaged part, $\langle \eta_i(t) \rangle$,
and a random fluctuation part, $\zeta_i(t)$,
%
\begin{equation}
  \eta_i(t) =
  \langle \eta_i(t) \rangle
  +
  \zeta_i(t)
  .
  \label{eq:h_split}
\end{equation}
%
The deterministic part can be related
to an ``ensemble average'',
where the ensemble consists of many similar simulation copies
sharing the same schedule $\alpha(t)$
and the same bias potential at time $t$.
%
The initial states and the stochastic forces
during the process may differ, however.

For sufficiently small $\alpha(t)$,
the bias potential remains roughly the same for a short period.
%
Then,
the sampling process is approximately Markovian, % of a finite order,
and we may assume Eq. \eqref{eq:pi_p_V}
for the deterministic part
%
\begin{align}
  \langle \eta_i(t) \rangle
  %&=
  %\frac{ \langle h_i(t) } { p_i } \rangle
  &\approx
  \frac{ \pi_i(t) } { p_i }
  =
  \frac{                       e^{-v_{*i}(t)} }
       { \sum_{j = 1}^n p_j \, e^{-v_{*j}(t)} }
  \notag
  \\
  &\approx
  1 - v_{*i}(t) + \sum_{j = 1}^n p_j \, v_{*j}(t)
  ,
  \notag
  %\label{eq:h_ave}
\end{align}
%
where we have assumed the smallness
of $v_{*i}(t)$ in the linear expansion.
%and used $\sum_{j=1}^n p_j = 1$
%on the second line.
%
\note{
The second step follows from
$$
\frac{ \langle h_i(t) \rangle }
     { p_i }
\approx
\frac{                       1 - v_{*i}  }
     { \sum_{ r = 1 }^n p_j (1 - v_{*j}) }
=
\frac{                       1 - v_{*i}  }
     { 1 - \sum_{ j = 1 }^n p_j \, v_{*j} }
.
$$
}%
%
Deducting the bin average, we get
%
\begin{align}
  \langle \eta_{*i}(t) \rangle
  \equiv
  \langle \eta_i(t) \rangle -
  \sum_{j=1}^n \rho_j \langle \eta_j(t) \rangle
  &\approx
  - v_{*i}(t)
  .
  \label{eq:sh_ave}
\end{align}

We will approximate the histogram fluctuation part, $\zeta_i(t)$,
as white noise,
and assume its validity for long simulations.
%
The bin-average-deducted fluctuation, $\zeta_{*i}(t)$,
%%
%\begin{equation}
%  \zeta_{*i}(t)
%  \equiv
%  \zeta_i(t)
%  -
%  \sum_{ j = 1 }^n
%    \rho_j \, \zeta_j(t)
%  ,
%  \notag
%  %\label{eq:xi_def}
%\end{equation}
%%
would also be governed by white noise as is $\zeta_i(t)$, i.e.,
\begin{equation}
  \langle \zeta_{*i}(t) \, \zeta_{*i}(t') \rangle
  = G_i \, \delta(t - t')
  ,
  \label{eq:G_def}
\end{equation}
for some $G_i$,
and $\delta(t)$ is the Dirac's delta function.
%
For non-white noise,
the equivalent value of $G_i$ can be approximately obtained as
$\sum_{t = -\infty}^\infty \langle \zeta_{*i}(t) \, \zeta_{*i}(0) \rangle$.
%
\note{More explicitly,
if the transition matrix of a period, $\tau$,
is $T^\tau_{ij}$,
then
$
  p_i p_j \langle \zeta_i(t) \, \zeta_j(t') \rangle
  =
  \langle h_i(t) \, h_j(t') \rangle
  -
  \langle h_i(t) \rangle \, \langle h_j(t') \rangle
  =
  T^{t - t'}_{ij} \, \pi_j - \pi_i \, \pi_j.
$
}%

From Eqs.
\eqref{eq:vt_diffeq}-\eqref{eq:sh_ave},
we get % a set of decoupled equations
%
\begin{equation}
  \dot v_{*i}(t)
  =
  \alpha(t) \, \left[ \langle \eta_{*i}(t) \rangle + \zeta_{*i}(t) \right]
  \approx
  -\alpha(t) \, \left[ v_{*i}(t) - \zeta_{*i}(t) \right]
  ,
  \notag
  %\label{eq:dxdt_singlebin}
\end{equation}
%
and the formal solution is % of Eq. \eqref{eq:dxdt_singlebin} is
%
\begin{equation}
  v_{*i}(T)
=
  v_{*i}(0) \, e^{-q(T)}
+
\int_0^T
  \dot y\bigl( q(t) \bigr) \, \zeta_{*i}(t) \, dt,
\notag
%\label{eq:xt_solution}
\end{equation}
%
where
%
$
q(t) \equiv \int_0^t \alpha(t') \, dt',
$
%
and
%
\begin{align}
y(q')
&\equiv
e^{q' - q(T)}.
\label{eq:y_def}
\end{align}

@@

The error depends implicitly on the curve, $\alpha(t)$,
or, equivalently, the curve, $q(t)$,
via $v_{*i}(T)$, and
our aim is to find the $q(t)$
that minimizes this error.
%
We now fix the endpoint value, $q(T)$,
and vary the curve, $q(t)$ ($0 < t < T$).
%
This is equivalent to varying the curve $y(q(t))$
with fixed endpoints,
$y\bigl( q(0) \bigr)  = e^{- q(T)}$
and
$y\bigl( q(T) \bigr) = 1$.
%
Note also that the first term on the right-hand side
of Eq. \eqref{eq:ET_average} is fixed during
the process,
while the second term is the action of a free particle.
Thus, the velocity of $y(q(t))$ should be a constant:
%
%We will further demand that
%%
%\begin{equation}
%  \lim_{t \to \infty} q(t) \to \infty.
%  \label{eq:qt_limit}
%\end{equation}
%%
%Then, for a large $t$,
%the second term on the right-hand side
%of Eq. \eqref{eq:xt_solution} can be neglected, and
%%
%\begin{align}
%\left\langle x^2(t) \right\rangle
%%&=
%%\int_0^t \int_0^t \dot y(t') \, \dot y(t'')
%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%%\notag
%%\\
%&=
%\int_0^t \int_0^t
%  \dot y\bigl( q(t') \bigr) \,
%  \dot y\bigl( q(t'') \bigr) \,
%  \kappa(t' - t'') \, dt'' \, dt'.
%\label{eq:x2t_average_asym}
%\end{align}
%
%By the Euler-Lagrange equation, we get
%$$
%\begin{aligned}
%0
%&=
%\frac{d}{dt} \int_0^T
%  \dot y\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
%\\
%&= \int_0^T
%  \ddot y\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
%\end{aligned}
%$$
%where we have dropped the boundary terms
%by assuming Eq. \eqref{eq:kappa_limit}
%and a large $T$.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t y^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
%In order for this to hold for any $t$,
%we must have
%%
%$
%\ddot y\bigl( q(t') \bigr) = 0
%,
%$
%or
\begin{equation}
  \dot y\bigl( q(t) \bigr) = c
  .
\label{eq:dydt_const}
\end{equation}
%
Then, by using Eq. \eqref{eq:y_def}, we have
%
\begin{equation}
  e^{ q(t) - q(T) }
  =
  c \, (t + t_0)
  =
  \frac{ t + t_0 }
       { T + t_0 }
  ,
\label{eq:expqt}
\end{equation}
where $t_0$ is some constant,
and the other constant, $c$, has been determined
from the $t = T$ case
in the second step.
%
Taking the logarithm,
and differentiating both sides
with respect to $t$,
we get the inverse-time schedule
%
\begin{equation}
  \alpha(t) = \frac{ 1 }{ t + t_0 }
  ,
\label{eq:alpha_invt1}
\end{equation}
%
and Eq. \eqref{eq:alpha_invt}
is the special case of $t_0 = 0$.
%

@@

The matrix, $\mathbf w$, is subject to three conditions.
%
First, we have a fixed-point condition\cite{bussi2006, dama2014}.
%
To sample the desired distribution,
$\mathbf p = (p_1, \dots, p_n)$,
the growth rate of the bias potential,
${\dot u}_i(t)$
in Eq. \eqref{eq:ut_diffeq_mbin},
should be a constant of $i$
if $h_j(t)$ were the same as $p_j$,
i.e.
$\sum_{j=1}^n w_{ij}$ should be a constant
to allow the bias potential to grow uniformly
in the asymptotic regime.
%
By a proper overall scaling of $\mathbf w$,
we may write this condition as
%
\begin{equation}
  \sum_{j = 1}^n w_{ij} = 1
  .
\label{eq:w_sumj}
\end{equation}
%
In other words, $(1, \dots, 1)^T$
is a right eigenvector of $\mathbf w$
with eigenvalue $1$.
%
Thus, the transpose $\mathbf w^T$
resembles a transition matrix,
except that some elements can be negative
in our cases.
%
Below we will take advantage of the analogy
and borrow techniques used
in studying transition matrices\cite{vankampen}
(we will only use results that are unaffected by the allowance
of negative elements into $\mathbf w$).

@@

We can further define
a generalized Fourier transform, $\mathcal{F}$,
for variable, $A_i$,
%
\begin{equation}
  {\tilde A}_k
  \equiv \mathcal{F}[A_i]_k
  \equiv \sum_{i = 1}^n \phi_{ki} \, A_i
  ,
  \notag
  %\label{eq:gft_def}
\end{equation}
%
and the inverse transform is
%
\begin{equation}
  A_i =
  \frac{1}{\rho_i} \sum_{k = 1}^n \phi_{ki} \, \tilde{A}_k
  .
  \notag
  %\label{eq:gft_inv}
\end{equation}
%
Using this definition,
we can rewrite Eq. \eqref{eq:ortho1} as
$\mathcal F[1]_k = \delta_{k0}$,
and for the bin average of any quantity, $f$, we have
$\bav{f}= \mathcal F[f_i]_0 = \tilde f_0$,
and
\begin{equation}
  \tilde f_{*k}
  = \tilde f_k - \bav{f} \, \delta_{k0}
  =
  \begin{dcases}
    0           & k = 0, \\
    \tilde f_k  & k > 0,
  \end{dcases}
  \label{eq:ftstar}
\end{equation}
which means that the Fourier transform of
a bin-average-deducted quantity always
has a vanishing first mode.

We can now use Eq. \eqref{eq:eig_w} to diagonalize
the multiple-bin updating scheme, Eq. \eqref{eq:mbin_update},
as
%
$$
{\tilde v}_k(t + 1) =
{\tilde v}_k(t) + \alpha(t) \, \lambda_k \,
{\tilde \eta}_k(t)
,
$$
or in the continuous-time form,
%
\begin{equation}
  \dot{\tilde v}_k(t)
  =
  \alpha(t) \, \lambda_k \, {\tilde \eta}_k(t)
  .
  \label{eq:ut_diffeq_mbin}
\end{equation}
%
Deducting the bin average yields
$$
  \dot {\tilde v}_{*k}(t)
  = \alpha(t) \, \lambda_k \, {\tilde \eta}_{*k}(t)
  \approx
  -\alpha(t) \, \lambda_k \,
  \bigl[ {\tilde v}_{*k}(t) - {\tilde \zeta}_{*k} (t) \bigr]
  ,
$$
where we have used the linear approximation
as in Sec. \ref{sec:sbin_diffeq}.

@@

The error defined in Eq. \eqref{eq:error_def}
can be rewritten in terms of the eigenmodes of $\mathbf w$,
%
\begin{align}
  \Err(T)
  =
  \sum_{i, j = 1}^n \sum_{k=0}^{n-1}
   \phi_{ki} \, \phi_{kj}
    \left\langle v_{*i}(T) \, v_{*j}(T) \right\rangle
  =
  \sum_{k = 0}^{n - 1}
    \left\langle
      {\tilde v}_{*k}^2(T)
    \right\rangle
  ,
\notag
\end{align}
%
where we have used Eq. \eqref{eq:eig_orthonormal_cols}.
%
By Eq. \eqref{eq:ftstar},
we can start the last sum %in Eq. \eqref{eq:xt2_sum}
from $k = 1$,
%
\begin{equation}
  \Err(T)
  =
  \sum_{k = 1}^{n - 1}
    \left\langle
      {\tilde v}_{*k}^2(T)
    \right\rangle
  .
\label{eq:xt2_sum}
\end{equation}
